Scalable analysis of genetic variation data in Python

### 2024-01-10 paper.tex
Add a temporary copy of Tim's diagram, + some refs

\section{Background}
% This current phrasing would annoy people, just spewing this
% out to roughly say what I think is relevant, we can refine later.

The study of genetics is currently transforming into a true big-data science.
Since the Human Genome Project, genomics has required working with
signicant volumes of data and the specialised methods
required to transform the raw biological signals into analysable data,
such as assembly[cite], mapping[cite] and variant calling[cite] are compute intensive.
However, beyond these initial steps required to generate the data, the actual
analysis of the genetic variation could typically be performed on a single
computer, and often with ad-hoc software written using shell utilities
and scripting languages such as Perl.
From the Human Genome Project, through 1000 Genomes, and now with
population scale genomics datasets such as UK Biobank, GeL, etc,
the requirements for analysing the data have far outgrown these
methods. Simple queries can require minutes to complete, which
has a negative effect on researcher productivity, as well as limiting
the analyses that can be performed. The financial cost of working
with such datasets is also substantial, which limits accessibility
to an elite set of well-funded groups.

@@

\begin{figure}
    \begin{tabular}{c}
    % The source for this figure is a google drawing
% Currently using
% https://docs.google.com/drawings/d/1NheB6LCvvkB4C0nAoSFwoYVZ3mtOPaseGmg_mZvcQ8I
% See https://github.com/pystatgen/sgkit-publication/issues/21 for
% discussion
\includegraphics[width=6cm]{diagrams/sgkit_xarray_diagram.pdf} \\
\includegraphics[width=6cm]{figures/data-scaling}
\end{tabular}
\caption{Chunked compressed storage of the genotype matrix using Zarr.
(A) Schematic of the chunked array storage used in sgkit
[Note: wrong fig. See
\url{https://github.com/pystatgen/sgkit-publication/issues/46} for
proposed update].
(B) Compression performance compared with BCF, savvy and genozip
based on simulated data (see text for details).
\label{fig-data-storage}}
\end{figure}

@@

\subsection{List of abbreviations}

>>>>

\section{Background}
[TODO rewrite this based on the first paragraph of the abstract above 
when cutting it down.]
The study of genetics is currently transforming into a true big-data science.
Since the Human Genome Project, genomics has required working with
signicant volumes of data and the specialised methods
required to transform the raw biological signals into analysable data,
such as assembly[cite], mapping[cite] and variant calling[cite] are compute intensive.
However, beyond these initial steps required to generate the data, the actual
analysis of the genetic variation could typically be performed on a single
computer, and often with ad-hoc software written using shell utilities
and scripting languages such as Perl.
From the Human Genome Project, through 1000 Genomes, and now with
population scale genomics datasets such as UK Biobank, GeL, etc,
the requirements for analysing the data have far outgrown these
methods. Simple queries can require minutes to complete, which
has a negative effect on researcher productivity, as well as limiting
the analyses that can be performed. The financial and 
environmental~\citep{grealey2022carbon,lannelongue2023greener} cost of working
with such datasets is also substantial.

@@

\begin{figure}
    \begin{tabular}{c}
% FIXME temporarily copying this in here until Tim's tikz code arrives.
% https://github.com/pystatgen/sgkit-publication/issues/46
\includegraphics[width=7cm]{diagrams/tmp-illustration.png} \\
\includegraphics[width=7cm]{figures/data-scaling}
\end{tabular}
\caption{Chunked compressed storage of the genotype matrix using Zarr.
(A) Schematic of the chunked array storage used in sgkit
[Note: wrong fig. See
\url{https://github.com/pystatgen/sgkit-publication/issues/46} for
proposed update].
(B) Compression performance compared with BCF, savvy and genozip
based on simulated data (see text for details).
\label{fig-data-storage}}
\end{figure}

@@

\subsection{List of abbreviations}
% If abbreviations are used in the text they should be defined in the text at
% first use, and a list of abbreviations should be provided in alphabetical
% order.

VCF: Variant Call Format;
